Project Code

### 2020-05-11 project2.Rmd
Nice working analysis

```{r libraries, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

@@

```{r}
p <- floor(seasonality.period/2)
weights <- rep(1/(2*p+1), times=2*p+1)
trend <- stats::filter(seis.h.ts, sides=2, filter = weights)
```

```{r show_overlap_and_diff, include=FALSE}
show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1) {
  overlap_plot <-
    autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
      geom_line(aes(y=base, col="red"), show.legend = F)
  span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
  diff <- compare-base
  diff_plot <-
    autoplot(diff, ylab = paste0("Difference (", diff.ylab, ")"), xlab = xlab) +
      labs(title = diff.title, subtitle = paste0(base.desc, " - ", compare.desc)) +
      scale_y_continuous(sec.axis = sec_axis(~./span, name=paste0("Difference (", diff.ylab.sec, ")"), labels = scales::percent_format(accuracy = diff.ylab.sec.accuracy))) +
      geom_hline(yintercept=mean(diff, na.rm=T), linetype="dashed", color = "red")
  return(list(overlap_plot, diff_plot))
}
```

@@

```{r}
f <- seasonality.period
weights.s <- c(0.5, rep(1,f-1), 0.5)/f
trend.s <- stats::filter(seis.h.ts, side=2, filter=weights.s)

plots <- show_overlap_and_diff(
  dec.addi$trend, trend.s, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("shift-filtered trend with f=", f),
  overlap.ylab = "Trend (nm)",
  diff.title = "Difference between trends",
  diff.ylab = "nm", diff.ylab.sec="% over trend span",
  diff.ylab.sec.accuracy = .001
)
plots[[1]]
plots[[2]]

@@

plots <- show_overlap_and_diff(
  dec.addi$figure, seasonality.figure, xlab="Hours",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Seasonality (nm)",
  diff.title = paste0("Difference between seasonalities (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over seasonality span"
)
plots[[1]]
plots[[2]]

@@

```{r}
residuals <- seis.h.ts - trend.s - seasonality

plots <- show_overlap_and_diff(
  dec.addi$random, residuals, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Residuals (nm)",
  diff.title = paste0("Difference between residuals (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over residuals span"
)
plots[[1]]
plots[[2]]

@@

```{r}
ts <- residuals+trend+seasonality

plots <- show_overlap_and_diff(
  seis.h.ts, ts, xlab="Weeks",
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span"
)
plots[[1]]
plots[[2]]

@@

```{r}
acf(residuals, na.action = na.omit)
pacf(residuals, na.action = na.omit)
seis.h.ts.ar <- Arima(seis.h.ts, c(1,0,0))
tsdiag(seis.h.ts.ar)
```

```{r}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- na.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
seis.h.msts %>% mstl() %>% autoplot() + xlab("Week")
seis.h.msts.mstl <- seis.h.msts %>% mstl()
plot(seis.h.msts.mstl[,2], main="trend")
lines(trend.s, col="red")
plot(rowSums(seis.h.msts.mstl[,3:4]), main="seasonality",type="l")
lines(c(rep(seasonality,seasons), seasonality[1:(((length(detrended)/168-seasons)*168))]), col="red")
plot(residuals,  main="residuals")
lines(seis.h.msts.mstl[,5], col="red")
autoplot(seis.h.msts.mstl[,5])
acf(seis.h.msts.mstl[,5], na.action = na.omit)
bptests <- c()
for (i in 1:10)
  bptests<- c(bptests, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
plot(bptests, ylim = c(0,1))
abline(h=0.05,col="blue",lty="dotted")

>>>>

```{r libraries, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(gridExtra) # for gg-plotting multiple plots
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

@@

```{r}
p <- floor(seasonality.period/2)
weights <- rep(1/(2*p+1), times=2*p+1)
trend <- stats::filter(na.aggregate(seis.h.ts, floor), sides=2, filter = weights)
```

```{r show_overlap_and_diff, include=FALSE}
show_overlap_and_diff <- function(base, compare, base.desc, compare.desc, xlab, overlap.ylab, diff.title, diff.ylab, diff.ylab.sec, diff.ylab.sec.accuracy=.1) {
  overlap_plot <-
    autoplot(compare, xlab=xlab, ylab=overlap.ylab) +
      geom_line(aes(y=base, col="red"), show.legend = F)
  span <- max(compare, base, na.rm = T) - min(compare, base, na.rm = T)
  diff <- compare-base
  accu.lims <- diff.ylab.sec.accuracy / 100 * span * c(-3,3)

  diff_plot <-
    autoplot(diff, ylab = paste0("Difference (", diff.ylab, ")"), xlab = xlab) +
      expand_limits(y=accu.lims) +
      labs(title = diff.title, subtitle = paste0(base.desc, " - ", compare.desc)) +
      scale_y_continuous(sec.axis = sec_axis(~./span, name=paste0("Difference (", diff.ylab.sec, ")"), labels = scales::percent_format(accuracy = diff.ylab.sec.accuracy))) +
      geom_hline(yintercept=mean(diff, na.rm=T), linetype="dashed", color = "red")
  return(list(overlap_plot, diff_plot))
}
```

@@

```{r}
f <- seasonality.period
weights.s <- c(0.5, rep(1,f-1), 0.5)/f
trend.s <- stats::filter(na.aggregate(seis.h.ts, floor), side=2, filter=weights.s)

plots <- show_overlap_and_diff(
  dec.addi$trend, trend.s, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("shift-filtered trend with f=", f),
  overlap.ylab = "Trend (nm)",
  diff.title = "Difference between trends",
  diff.ylab = "nm", diff.ylab.sec="% over trend span",
  diff.ylab.sec.accuracy = .01
)
plots[[1]]
plots[[2]]

@@

plots <- show_overlap_and_diff(
  dec.addi$figure, seasonality.figure, xlab="Hours",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Seasonality (nm)",
  diff.title = paste0("Difference between seasonalities (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over seasonality span",
  diff.ylab.sec.accuracy = .01
)
plots[[1]]
plots[[2]]

@@

```{r}
residuals <- na.aggregate(seis.h.ts, floor) - trend.s - seasonality

plots <- show_overlap_and_diff(
  dec.addi$random, residuals, xlab="Weeks",
  base.desc = "additive decomposition",
  compare.desc = paste0("mean-over-period after removal of shift-filtered trend with f=", f),
  overlap.ylab = "Residuals (nm)",
  diff.title = paste0("Difference between residuals (f=", seasonality.period, ")"),
  diff.ylab = "nm", diff.ylab.sec="% over residuals span",
  diff.ylab.sec.accuracy = .01
)
plots[[1]]
plots[[2]]

@@

```{r}
ts <- residuals+trend.s+seasonality

plots <- show_overlap_and_diff(
  seis.h.ts, ts, xlab="Weeks",
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plots[[1]]
plots[[2]]

plots <- show_overlap_and_diff(
  seis.h.ts, dec.addi$trend+dec.addi$seasonal+dec.addi$random, xlab="Weeks",
  base.desc = "original series",
  compare.desc = "reconstructed series",
  overlap.ylab = "Mean hourly movement (nm)",
  diff.title = "Difference between ts",
  diff.ylab = "nm", diff.ylab.sec="% over ts span",
  diff.ylab.sec.accuracy = .01
)
plots[[1]]
plots[[2]]

@@

```{r residuals cf arima, fig.height=960, fig.width=960, fig.path="/tmp/"}
residuals.acf <- autoplot(acf(residuals, na.action = na.pass, plot=F))
residuals.pacf <- autoplot(pacf(residuals, na.action = na.pass, plot=F))
plot_grid(residuals.acf, residuals.pacf, labels = "AUTO", nrow=2)

seis.h.ts.ar <- Arima(seis.h.ts, c(1,0,0))
tsdiag(seis.h.ts.ar)
library(cowplot)
install.packages("cowplot")
```

```{r}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- na.aggregate(build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end, force_ms = T))
seis.h.msts %>% mstl() %>% autoplot() + xlab("Week")
seis.h.msts.mstl <- seis.h.msts %>% mstl()
plot(seis.h.msts.mstl[,2], main="trend")
lines(trend.s, col="red")
plot(rowSums(seis.h.msts.mstl[,3:4]), main="seasonality",type="l")
lines(seasonality, col="red")
plot(residuals,  main="residuals")
lines(seis.h.msts.mstl[,5], col="red")
autoplot(seis.h.msts.mstl[,5])
acf(seis.h.msts.mstl[,5], na.action = na.omit)
bptests <- c()
for (i in 1:10)
  bptests<- c(bptests, Box.test(seis.h.msts.mstl[,5], lag = i, fitdf = 0, type = "Box-Pierce")$p.value)
plot(bptests, ylim = c(0,1))
abline(h=0.05,col="blue",lty="dotted")
